Estimating initial state and quark-gluon plasma medium properties \\ using a hybrid model with nucleon substructure \\ calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV

### 2018-08-01 paper/substructure.tex
Some extra work on the summary

Lastly, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of order $v_m$ and $v_n$ \cite{Bilandzic:2013kga, ALICE:2016kpq},
\begin{align}
  \mathrm{SC}(m, n) &= \langle\langle \cos[m(\phi_1 - \phi_3) + n(\phi_2-\phi_4)]\rangle\rangle \nonumber \\
\nonumber &- \langle\langle\cos[m(\phi_1-\phi_2)]\rangle\rangle\langle\langle\cos[n(\phi_1-\phi_2)]\rangle\rangle \label{eq:scmn}\\
&\approx \langle v_m^2 v_n^2 \rangle - \langle v_m^2\rangle\langle v_n^2\rangle.
\end{align}
We show these model prediction in Fig.~\ref{fig:flow_corr} along with the \emph{normalized} symmetric cumulants
\begin{equation}
  \mathrm{NSC}(m,n) = SC(m,n)/\langle v_m^2\rangle\langle v_n^2\rangle,
\end{equation}
which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
The solid lines are model calculations using the MAP parameters from the present analysis and the dashed lines are model calculations using the MAP parameters of Ref.~\note\ which did not include nucleon substructure.
The gap between $\mathrm{SC}(3,2)$ and $\mathrm{SC}(4,2)$ is generally wider in the present analysis, as is the gap between the normalized symmetric cumulants $\mathrm{NSC}(3,2)$ and $\mathrm{NSC}(4,2)$.
This may alleviate a common problem

\begin{figure}
  \includegraphics{flow_corr}
  \caption{Model calculations of the symmetric cumulants (top) and normalized symmetric cumulants (bottom) for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters. The solid lines are the MAP estimate of the present analysis (with nucleon substructure), and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} (without nucleon substructure). In general, most model parameters are different between the two studies.}
  \label{fig:flow_corr}
\end{figure}

@@

In this work, we performed a semi-exhaustive search for a unified description of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using Bayesian methods to rigorously calibrate and constrain free parameters of a flexible nuclear collision model based on viscous hydrodynamics.
The goal of our study was two fold.
First, we aimed to establish whether or not our hydrodynamic framework was able to describe $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV simultaneously, i.e.\ we attempted to falsify the framework.
And second, given that the former is true, we wished to obtain estimates for the \emph{true} model parameters given the assumptions of our model and the evidence provided by the experimental data.

Perhaps the most important difference between the present study and previous Bayesian analyses of heavy-ion collision models \note\ is the inclusion of variable nucleon substructure.
We described the QGP initial state using a modified version of the parametric \trento\ model, with new parameters to vary the fluctuating size and shape of each nucleon.
The nucleon was modeled as a cluster of $n_c$ Gaussian constituents (hot spots), and each constituent was described by a Gaussian density profile of width $v$.
We also varied a parameter $w$ which controlled the sampling radius of the constituent positions inside each nucleon and hence the resulting RMS width of the ensemble averaged nucleon density.

The transport dynamics were simulated using an event-by-event multi-stage nuclear collision model described by a pre-equilibrium free-streaming stage coupled to boost-invariant viscous hydrodynamics for the early hot and dense regions of the collision, and a microscopic hadronic afterburner to simulate hadronic interactions .
We parametrized various sources of uncertainty including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the switching or particlization temperature where the medium description switches from hydrodynamics to microscopic Boltzmann transport.

%Unfortunately, validating (or invalidating) a hydrodynamic picture of $p$-Pb and Pb-Pb collisions is not as simple as comparing a single model calculation to experimental data.
%Hydrodynamic simulations typically include a number of free parameters which quantify uncertainties in the initial conditions and QGP medium, and these parameters can dramatically affect model predictions.
%It's thus imperative to test all possible combinations of parameters in order to compare the full landscape of physically reasonable model calculations to experimental data.
%If all possible combinations of model parameters fail to describe both collision systems simultaneously, then it's probably safe to say that the underlying assumptions of the chosen framework are either incomplete or invalid.
%Similarly, if a single set of model parameters describes $p$-Pb and Pb-Pb bulk observables simultaneously, then it's compelling evidence that the hydrodynamic framework generalizes to small collision systems.
%
%There is also much more that can be learned about a model by testing all possible parameter combinations as opposed to a single set of values.
%If the model works exceptionally well, but only for a certain range of parameter values, it's evidence that those parameter values are somehow constrained by the data.
%Moreover, rigorous statistical methods may then be used to report parameter estimates with meaningful uncertainties which account for the accuracy of the chosen framework and the resolving power of the experimental data.

>>>>

Lastly, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of different order \cite{Bilandzic:2013kga, ALICE:2016kpq},
\begin{align}
  \mathrm{SC}(m, n) &= \langle\langle \cos[m(\phi_1 - \phi_3) + n(\phi_2-\phi_4)]\rangle\rangle \nonumber \\
\nonumber &- \langle\langle\cos[m(\phi_1-\phi_2)]\rangle\rangle\langle\langle\cos[n(\phi_1-\phi_2)]\rangle\rangle \label{eq:scmn}\\
&\approx \langle v_m^2 v_n^2 \rangle - \langle v_m^2\rangle\langle v_n^2\rangle.
\end{align}
We show these model prediction in Fig.~\ref{fig:flow_corr} along with the \emph{normalized} symmetric cumulants
\begin{equation}
  \mathrm{NSC}(m,n) = SC(m,n)/\langle v_m^2\rangle\langle v_n^2\rangle,
\end{equation}
which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
The solid lines are the MAP estimate of the present study and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} which did not include nucleon substructure and was calibrated on Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV.
We observe that the gap between $\mathrm{SC}(3,2)$ and $\mathrm{SC}(4,2)$ is generally wider in the present analysis than in Ref.~\cite{Bernhard:2018hnz}, as is the gap between the normalized symmetric cumulants $\mathrm{NSC}(3,2)$ and $\mathrm{NSC}(4,2)$.

We emphasize that multiple aspects of the two analyses are different such as the collision systems and beam energies considered, the observables which were included in each calibration, and the existence of nucleon substructure to name a few.
Thus we can only speculate what might have lead to the large difference in the symmetric flow cumulants.
Two reasonable culprits would be the inclusion of nucleon substructure and the large difference in the preferred pre-equilibrium free-streaming time determined by the two studies.
Identifying the source of the effect would be an interesting study for a future project.

\begin{figure}
  \includegraphics{flow_corr}
  \caption{Model calculations of the symmetric cumulants (top) and normalized symmetric cumulants (bottom) for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters. The solid lines are the MAP estimate of the present analysis (with nucleon substructure), and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} (without nucleon substructure) which was calibrated on Pb-Pb observables at $\sqrts=2.76$ and 5.02~TeV. In general, most model parameters are somewhat different between the two studies.}
  \label{fig:flow_corr}
\end{figure}

@@

In this work, we performed a semi-exhaustive search for a unified description of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using Bayesian methods to rigorously calibrate and constrain free parameters of a flexible nuclear collision model based on viscous hydrodynamics.
The goal of our study was two fold.
First, we aimed to establish whether or not our hydrodynamic framework was able to describe both collision systems simultaneously.
And second, in the event that the former was true, we wished to obtain estimates for the \emph{true} parameters of our model given the assumptions of our framework and the evidence provided by the experimental data.

We built, for this purpose, a flexible multi-stage nuclear collision model characterized by a number of free parameters which vary theoretically uncertain aspects of the framework such as the QGP initial conditions and hydrodynamic transport properties.
For the QGP initial conditions, we developed a modified version of the \trento\ model which adds new parameters to vary the fluctuating size and shape of each nucleon.
Specifically, we modeled each nucleon as a cluster of $n_c$ constituents (hot spots), where each constituent is described by a Gaussian density profile of width $v$.
We also introduced a parameter $w$ which effectively varies the width of the Gaussian distribution used to sample constituent positions inside each nucleon.

The transport dynamics of the collision were simulated using a pre-equilibrium free-streaming stage of variable duration followed by boost-invariant viscous hydrodynamics for hot and dense regions of the fireball and a microscopic hadronic afterburner for the dilute corona.
We parametrized various sources of uncertainty in each stage of the collision including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the isothermal particlization temperature used to switch from a hydrodynamic description to microscopic Boltzmann transport.

We applied Bayesian parameter estimation which were developed to estimate the parameters of computationally expensive model.

\medskip

...work in progress

\vfill
